---
title: "Hands-on Exercise 1"
author: "Michael Djohan"
date: "`r Sys.Date()`"
---

## **1. Getting Started**

### Install and launching R packages.

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

### Importing the data

```{r}
#| output: false
exam_data <- read_csv("data/Exam_data.csv")
```

## 2. Try-It-Yourself Exercise

### Working with facet

*Example 1 Plotting trellis boxplot using facet_wrap by GENDER*

```{r}
ggplot(data = exam_data,
       aes(x = CLASS,
           y = MATHS)) +
  geom_boxplot() +
  facet_wrap(~ GENDER)
```

*Example 2 Plotting trellis boxplot using facet_grid by GENDER*

```{r}
ggplot(data = exam_data,
       aes(x = CLASS,
           y = MATHS)) +
  geom_boxplot() +
  facet_grid(GENDER ~ .)
```

*Example 3 Plotting 2-D trellis boxplot using facet_grid*

```{r}
ggplot(data = exam_data,
       aes(x = GENDER,
           y = MATHS)) +
  geom_boxplot() +
  facet_grid(GENDER ~ CLASS)
```

### Working with theme

*Example 4 Changing theme of bar plot*

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme(panel.background = element_rect(fill = "light blue"),
        panel.grid = element_line(color = "white"))
```

### Designing Data-driven Graphics for Analysis

*Exercise 1*

```{r}
ggplot(data=exam_data, 
       aes(x = fct_infreq(RACE))) +
  geom_bar() +
  geom_text(stat = "count", 
            aes(label = paste0(after_stat(count), ", ",
                        round(after_stat(count)/sum(after_stat(count))*100,
                        1), "%")), 
            vjust = -1) +
  labs(
    x = "Race",
    y = "No. of\nPupils") +
  theme(axis.title.y = element_text(angle = 0))
```

*Exercise 2*

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 color = "black",
                 fill = "light blue") +
  geom_vline(aes(xintercept = mean(MATHS)),
             linetype = "dashed",
             color = "red") +
  geom_vline(aes(xintercept = median(MATHS)),
             linetype = "dashed",
             color = "black")
```

*Exercise 3*

```{r}
ggplot(data = exam_data,
       aes(x = ENGLISH)) +
  geom_histogram(data = exam_data |> select(ENGLISH), 
                 bins = 30,
                 alpha = 0.3) +
  geom_histogram(aes(fill = GENDER),
                 bins = 30,
                 color = "black",
                 show.legend = FALSE) +
  facet_grid(~ GENDER) +
  theme_bw()
```

*Exercise 4*

```{r}
ggplot(data = exam_data,
       aes(y = ENGLISH,
           x = MATHS)) +
  geom_point() +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  geom_vline(aes(xintercept = 50),
             linetype = "dashed",
             color = "grey40") +
  geom_hline(aes(yintercept = 50),
             linetype = "dashed",
             color = "grey40") 
```
