---
title: "Hands-on Exercise 5: Visual Multivariate Analysis"
author: "Michael Djohan"
date: 08 February 2023
date-modified: "`r Sys.Date()`"
fontsize: smaller
execute: 
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

## 1. Plotting Ternary Diagram with R

### Install and launching R packages.

The code chunk below uses p_load() of pacman package to check if packages are installed in the computer. If they are, then they will be launched into R. The R packages installed are:

-   [**ggtern**](http://www.ggtern.com/), a ggplot extension specially designed to plot ternary diagrams. The package will be used to plot static ternary plots.
-   [**Plotly R**](https://plot.ly/r/), an R package for creating interactive web-based graphs via plotly's JavaScript graphing library, plotly.js . The **plotly R** libary contains the *ggplotly* function, which will convert **ggplot2** figures into a Plotly object.

```{r}
pacman::p_load(ggtern, plotly, tidyverse)
```

### Importing the data

```{r}
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv") 
```

### Preparing the data

```{r}
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8]))%>%
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%
  filter(TOTAL > 0)

agpop_mutated
```
